---
title: "Lodes OD Plots"
author: "Owen Hart"
date: "7/15/2020"
output: html_document
---

```{r}
library(R.utils)
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(sf)
library(leaflet)
library(mapview)
library(reshape2)
library(raster)
library(tigris)
```


reading in filtered lodes data with geometries
```{r}
wasco17_od <- readRDS("../data/lodes/filtered/wasco17_od.Rds")
wasco16_od <- readRDS("../data/lodes/filtered/wasco16_od.Rds")
wasco15_od <- readRDS("../data/lodes/filtered/wasco17_od.Rds")
```

```{r}
# select tracts from geocode strings
wasco17_od$w_geocode_tract <- substr(wasco17_od$w_geocode, start = 1, stop = 11)
wasco17_od$h_geocode_tract <- substr(wasco17_od$h_geocode, start = 1, stop = 11)

wasco16_od$w_geocode_tract <- substr(wasco16_od$w_geocode, start = 1, stop = 11)
wasco16_od$h_geocode_tract <- substr(wasco16_od$h_geocode, start = 1, stop = 11)

wasco15_od$w_geocode_tract <- substr(wasco15_od$w_geocode, start = 1, stop = 11)
wasco15_od$h_geocode_tract <- substr(wasco15_od$h_geocode, start = 1, stop = 11)
```

aggregate by tract level
```{r}
jobs_all_2017 <- wasco17_od%>%
  group_by(createdate, w_geocode_tract, h_geocode_tract) %>% 
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
jobs_all_2016 <- wasco16_od %>%
  group_by(createdate, w_geocode_tract, h_geocode_tract) %>% 
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
jobs_all_2015 <- wasco15_od %>%
  group_by(createdate, w_geocode_tract, h_geocode_tract) %>% 
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
```

```{r}
jobs_all_2017$h_state <- substr(jobs_all_2017$h_geocode_tract, 1, 2)
jobs_all_2017$h_county <- substr(jobs_all_2017$h_geocode_tract, 3, 5)
jobs_all_2017$h_tract <- substr(jobs_all_2017$h_geocode_tract, 6, 11)

jobs_all_2016$h_state <- substr(jobs_all_2016$h_geocode_tract, 1, 2)
jobs_all_2016$h_county <- substr(jobs_all_2016$h_geocode_tract, 3, 5)
jobs_all_2016$h_tract <- substr(jobs_all_2016$h_geocode_tract, 6, 11)

jobs_all_2015$h_state <- substr(jobs_all_2015$h_geocode_tract, 1, 2)
jobs_all_2015$h_county <- substr(jobs_all_2015$h_geocode_tract, 3, 5)
jobs_all_2015$h_tract <- substr(jobs_all_2015$h_geocode_tract, 6, 11)

#now work geoid
jobs_all_2017$w_state <- substr(jobs_all_2017$w_geocode_tract, 1, 2)
jobs_all_2017$w_county <- substr(jobs_all_2017$w_geocode_tract, 3, 5)
jobs_all_2017$w_tract <- substr(jobs_all_2017$w_geocode_tract, 6, 11)

jobs_all_2016$w_state <- substr(jobs_all_2016$w_geocode_tract, 1, 2)
jobs_all_2016$w_county <- substr(jobs_all_2016$w_geocode_tract, 3, 5)
jobs_all_2016$w_tract <- substr(jobs_all_2016$w_geocode_tract, 6, 11)

jobs_all_2015$w_state <- substr(jobs_all_2015$w_geocode_tract, 1, 2)
jobs_all_2015$w_county <- substr(jobs_all_2015$w_geocode_tract, 3, 5)
jobs_all_2015$w_tract <- substr(jobs_all_2015$w_geocode_tract, 6, 11)
```

Must load in Oregon XWalk csv
```{r}
#download.file(
#  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/or_xwalk.csv.gz", 
#  destfile = "../data/lodes/or_xwalk.csv.gz")
#gunzip("../data/lodes/or_xwalk.csv.gz")
or_xw <- read_csv("../data/lodes/or_xwalk.csv", col_types = cols(cty = col_character()))
or_xw <- or_xw %>% filter(st == 41)
```
What % of jobs flow from outside of Fairfax County?
```{r}
bool_cond17 <-jobs_all_2017$h_county != "065" & (jobs_all_2017$w_county == "065" & jobs_all_2017$w_state == "41")
bool_cond16 <-jobs_all_2016$h_county != "065" & (jobs_all_2016$w_county == "065" & jobs_all_2016$w_state == "41")
bool_cond15 <-jobs_all_2015$h_county != "065" & (jobs_all_2015$w_county == "065" & jobs_all_2015$w_state == "41")
flows_in_17 <-jobs_all_2017[bool_cond17, ]
flows_in_16 <-jobs_all_2016[bool_cond16, ]
flows_in_15 <-jobs_all_2015[bool_cond15, ]
```

Of the flows into Wasco County from outside, what percent are from other counties in Oregon/Washington?
```{r}
#2017
agg_cnty_OR_2017 <- flows_in_17[flows_in_17$h_state == "41", ] %>%
  group_by(h_county) %>%
  summarise(S000 = sum(S000))
#Use OR crosswalk to get county names
or_xw$h_county <- substring(or_xw$cty, 3, 6)
essential <- or_xw %>%
  select(h_county, ctyname)
agg_cnty_OR_with_names_2017 <- inner_join(x = agg_cnty_OR_2017, y = unique(essential), by = "h_county")

trimmed_2017 <- agg_cnty_OR_with_names_2017[order(-agg_cnty_OR_with_names_2017$S000), ][1:10, ]

#2016
agg_cnty_OR_2016 <- flows_in_16[flows_in_16$h_state == "41", ] %>%
  group_by(h_county) %>%
  summarise(S000 = sum(S000))
agg_cnty_OR_with_names_2016 <- inner_join(x = agg_cnty_OR_2016, y = unique(essential), by = "h_county")
trimmed_2016 <- agg_cnty_OR_with_names_2016[order(-agg_cnty_OR_with_names_2016$S000), ][1:10, ]


#2015
agg_cnty_OR_2015 <- flows_in_15[flows_in_15$h_state == "41", ] %>%
  group_by(h_county) %>%
  summarise(S000 = sum(S000))
agg_cnty_OR_with_names_2015 <- inner_join(x = agg_cnty_OR_2015, y = unique(essential), by = "h_county")
trimmed_2015 <- agg_cnty_OR_with_names_2015[order(-agg_cnty_OR_with_names_2015$S000), ][1:10, ]
```
Aggregating by county over time (line plot)
```{r}
top_10_in_over_time <- inner_join(trimmed_2017, agg_cnty_OR_with_names_2016, by = "ctyname")
top_10_in_over_time <- top_10_in_over_time %>%
  select(h_county.x, ctyname, S000.x, S000.y)
top_10_in_over_time$'2017' <- top_10_in_over_time$S000.x
top_10_in_over_time$'2016' <- top_10_in_over_time$S000.y
top_10_in_over_time <- inner_join(top_10_in_over_time, agg_cnty_OR_with_names_2015, by = "ctyname") %>%
  select(h_county.x, ctyname, '2017', '2016', S000)
top_10_in_over_time$'2015' <- top_10_in_over_time$S000
top_10_in_over_time$h_county <- top_10_in_over_time$h_county.x
top_10_in_over_time <- top_10_in_over_time %>% select(-S000, -h_county.x)
top_10_in_over_time <- top_10_in_over_time %>% select(ctyname, '2017', '2016', '2015')
top_10_in_trans <- data.frame(t(top_10_in_over_time[, -1]))
colnames(top_10_in_trans) <- top_10_in_over_time$ctyname
top_10_in_trans$year <- c(2017, 2016, 2015) #SAVE NOW
write_csv(top_10_in_trans, "../data/app_10_inflows_wasco.csv")
#line plot for all jobs 2015-2017 by county
ggplot(top_10_in_trans, aes(x = year)) +
  ggtitle("Number of jobs flowing into Wasco County\nfrom other counties in Oregon \n2015-2017") +
  labs(x = "Year", y = "Number of Jobs", colour = "County") + 
  geom_line(aes(y = `Hood River County, OR`, color = "Hood River County, OR")) + 
  geom_line(aes(y = `Multnomah County, OR`, color = "Multnomah County, OR")) + 
  geom_line(aes(y = `Clackamas County, OR`, color = "Clackamas County, OR")) + 
  geom_line(aes(y = `Marion County, OR`, color = "Marion County, OR")) + 
  geom_line(aes(y = `Washington County, OR`, color = "Washington County, OR")) + 
  geom_line(aes(y = `Deschutes County, OR`, color = "Deschutes County, OR")) + 
  geom_line(aes(y = `Jefferson County, OR`, color = "Jefferson County, OR")) + 
  geom_line(aes(y = `Lane County, OR`, color = "Lane County, OR")) + 
  geom_line(aes(y = `Umatilla County, OR`, color = "Umatilla County, OR")) + 
  geom_line(aes(y = `Sherman County, OR`, color = "Sherman County, OR")) +
  scale_x_continuous(breaks = 0:2100)
ggsave(path = "../output/lodes_maps/flows/in_flows", device = "png", filename = "in_county_flows_all_years.png", plot = last_plot())
```

Outflows (those who live in wasco county but do not work there, open to other states)
```{r}
bool_cond17 <- jobs_all_2017$h_county == "065" & jobs_all_2017$w_county != "065"
bool_cond16 <- jobs_all_2016$h_county == "065" & jobs_all_2016$w_county != "065"
bool_cond15 <- jobs_all_2015$h_county == "065" & jobs_all_2015$w_county != "065"

flows_out_17 <-jobs_all_2017[bool_cond17, ]
flows_out_16 <-jobs_all_2016[bool_cond16, ]
flows_out_15 <-jobs_all_2015[bool_cond15, ]
```

Most popular counties that those who live in Wasco County but not work in, go to work
```{r}
out_agg_2017 <- flows_out_17 %>%
  group_by(w_county) %>%
  summarise(S000 = sum(S000))
or_xw$w_county <- substring(or_xw$cty, 3, 6)
essential <- or_xw %>%
  select(w_county, ctyname)
agg_out_names_2017 <- inner_join(x = out_agg_2017, y = unique(essential), by = "w_county")
trimmed_out_2017 <- agg_out_names_2017[order(-agg_out_names_2017$S000), ][1:10, ]

out_agg_2016 <- flows_out_16 %>%
  group_by(w_county) %>%
  summarise(S000 = sum(S000))
agg_out_names_2016 <- inner_join(x = out_agg_2016, y = unique(essential), by = "w_county")
trimmed_out_2016 <- agg_out_names_2016[order(-agg_out_names_2016$S000), ][1:10, ]

out_agg_2015 <- flows_out_15 %>%
  group_by(w_county) %>%
  summarise(S000 = sum(S000))
agg_out_names_2015 <- inner_join(x = out_agg_2015, y = unique(essential), by = "w_county")
trimmed_out_2015 <- agg_out_names_2015[order(-agg_out_names_2015$S000), ][1:10, ]
```

outflows over time!!
```{r}
top_10_out_over_time <- inner_join(trimmed_out_2017, agg_out_names_2016, by = "ctyname")
top_10_out_over_time <- top_10_out_over_time %>%
  select(w_county.x, ctyname, S000.x, S000.y)
top_10_out_over_time$'2017' <- top_10_out_over_time$S000.x
top_10_out_over_time$'2016' <- top_10_out_over_time$S000.y
top_10_out_over_time <- inner_join(top_10_out_over_time, agg_cnty_OR_with_names_2015, by = "ctyname") %>%
  select(w_county.x, ctyname, '2017', '2016', S000)
top_10_out_over_time$'2015' <- top_10_out_over_time$S000
top_10_out_over_time$h_county <- top_10_out_over_time$w_county.x
top_10_out_over_time <- top_10_out_over_time %>% select(-S000, -w_county.x)
top_10_out_over_time <- top_10_out_over_time %>% select(ctyname, '2017', '2016', '2015')
top_10_out_trans <- data.frame(t(top_10_out_over_time[, -1]))
colnames(top_10_out_trans) <- top_10_out_over_time$ctyname
top_10_out_trans$year <- c(2017, 2016, 2015)
write_csv(top_10_out_trans, "../data/app_10_outflows_wasco.csv")

#line plot for all jobs 2015-2017 by county
ggplot(top_10_out_trans, aes(x = year)) +
  ggtitle("Number of jobs flowing from Wasco County\ninto other counties in Oregon \n2015-2017") +
  labs(x = "Year", y = "Number of Jobs", colour = "County") + 
  geom_line(aes(y = `Hood River County, OR`, color = "Hood River County, OR")) + 
  geom_line(aes(y = `Multnomah County, OR`, color = "Multnomah County, OR")) + 
  geom_line(aes(y = `Clackamas County, OR`, color = "Clackamas County, OR")) + 
  geom_line(aes(y = `Deschutes County, OR`, color = "Deschutes County, OR")) + 
  geom_line(aes(y = `Washington County, OR`, color = "Washington County, OR")) + 
  geom_line(aes(y = `Marion County, OR`, color = "Marion County, OR")) + 
  geom_line(aes(y = `Jefferson County, OR`, color = "Jefferson County, OR")) + 
  geom_line(aes(y = `Umatilla County, OR`, color = "Umatilla County, OR")) + 
  geom_line(aes(y = `Lane County, OR`, color = "Lane County, OR")) + 
  geom_line(aes(y = `Sherman County, OR`, color = "Sherman County, OR")) +
  scale_x_continuous(breaks = 0:2100)

ggsave(path = "../output/lodes_maps/flows/out_flows", device = "png", filename = "out_county_flows_all_years.png", plot = last_plot())
```





Mapping S. Wasco
```{r}
wasco_points <- (blocks("OR", county = "Wasco"))
wasco_lines <- data.frame(wasco_points)
south_wasco_points <- st_read("../data/shps/swsd")
```

Reading in filtered file again
```{r}
wasco17_od2 <- readRDS("../data/lodes/filtered/wasco17_od.Rds")
wasco16_od2 <- readRDS("../data/lodes/filtered/wasco16_od.Rds")
wasco15_od2 <- readRDS("../data/lodes/filtered/wasco17_od.Rds")
```

Joining with Wasco Lines
```{r}
wascolines_17 <- inner_join(wasco17_od2, wasco_lines, by = c("w_geocode" = "GEOID10"))
wascolines_16 <- inner_join(wasco16_od2, wasco_lines, by = c("w_geocode" = "GEOID10"))
wascolines_15 <- inner_join(wasco15_od2, wasco_lines, by = c("w_geocode" = "GEOID10"))
```

Grouping by work geocode (because that is going to be Wasco County)
```{r}
agg_17 <- wascolines_17 %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10, geometry) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
saveRDS(st_as_sf(agg_17), "../data/app_lodes_od_agg_2017.Rds")
agg_16 <- wascolines_16 %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10, geometry) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
saveRDS(st_as_sf(agg_16), "../data/app_lodes_od_agg_2016.Rds")
agg_15 <- wascolines_15 %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10, geometry) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
saveRDS(st_as_sf(agg_15), "../data/app_lodes_od_agg_2015.Rds")
```



interactive maps

```{r}
#S000 (all jobs) by year
qtileS000 <- colorQuantile(c('#D1E0BF', '#E57200'), agg_17$S000, 5)
od_S000leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 2,
    opacity = 1,
    group = "Basemap",
    label = "South Wasco Region")

od_S000leaf <- od_S000leaf %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2017",
    fillColor = ~qtileS000(agg_17$S000),
    label = agg_17$S000) %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2016",
    fillColor = ~qtileS000(agg_16$S000),
    label = agg_16$S000) %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2015",
    fillColor = ~qtileS000(agg_15$S000),
    label = agg_15$S000) %>%
  addLegend(
    data = agg_17,
    "bottomright",
    pal = qtileS000,
    values = ~ S000,
    title = "Wasco County All Job Density",
    opacity = 1,
    na.label = "NA") %>%
  addLayersControl(
#    baseGroups = c("South Wasco School District"),
    baseGroups = c("2017", "2016", "2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("2016", "2015"))
```

```{r}
#SI01 (Goods Producing industry sectors) by year
qtileS000 <- colorQuantile(c('#D1E0BF', '#E57200'), agg_17$S000, 5)
od_SI01leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 2,
    opacity = 1,
    group = "Basemap",
    label = "South Wasco Region")

od_SI01leaf <- od_SI01leaf %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2017",
    fillColor = ~qtileS000(agg_17$SI01),
    label = agg_17$SI01) %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2016",
    fillColor = ~qtileS000(agg_16$SI01),
    label = agg_16$SI01) %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2015",
    fillColor = ~qtileS000(agg_15$SI01),
    label = agg_15$SI01) %>%
  addLegend(
    data = agg_17,
    "bottomright",
    pal = qtileS000,
    values = ~ S000,
    title = "Goods Producing Industry\nJob Density",
    opacity = 1,
    na.label = "NA") %>%
  addLayersControl(
#    baseGroups = c("South Wasco School District"),
    baseGroups = c("2017", "2016", "2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("2016", "2015"))
```


```{r}
#SI02 (Trade, Transportation, and Utilities industry sectors) by year
qtileS000 <- colorQuantile(c('#D1E0BF', '#E57200'), agg_17$S000, 5)
od_SI02leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 2,
    opacity = 1,
    group = "Basemap",
    label = "South Wasco Region")

od_SI02leaf <- od_SI02leaf %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2017",
    fillColor = ~qtileS000(agg_17$SI02),
    label = agg_17$SI02) %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2016",
    fillColor = ~qtileS000(agg_16$SI02),
    label = agg_16$SI02) %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2015",
    fillColor = ~qtileS000(agg_15$SI02),
    label = agg_15$SI02) %>%
  addLegend(
    data = agg_17,
    "bottomright",
    pal = qtileS000,
    values = ~ S000,
    title = "Trade, Transportation,\nand Utilities Industry\nJob Density",
    opacity = 1,
    na.label = "NA") %>%
  addLayersControl(
#    baseGroups = c("South Wasco School District"),
    baseGroups = c("2017", "2016", "2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("2016", "2015"))
```
```{r}
#SI03 (All Other Services industry sectors) by year
qtileS000 <- colorQuantile(c('#D1E0BF', '#E57200'), agg_17$S000, 5)
od_SI03leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 2,
    opacity = 1,
    group = "Basemap",
    label = "South Wasco Region")

od_SI03leaf <- od_SI03leaf %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2017",
    fillColor = ~qtileS000(agg_17$SI03),
    label = agg_17$SI03) %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2016",
    fillColor = ~qtileS000(agg_16$SI03),
    label = agg_16$SI03) %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "2015",
    fillColor = ~qtileS000(agg_15$SI03),
    label = agg_15$SI03) %>%
  addLegend(
    data = agg_17,
    "bottomright",
    pal = qtileS000,
    values = ~ S000,
    title = "All Other Services Industry\nJob Density",
    opacity = 1,
    na.label = "NA") %>%
  addLayersControl(
#    baseGroups = c("South Wasco School District"),
    baseGroups = c("2017", "2016", "2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("2016", "2015"))
```


Showing all maps
```{r}
od_S000leaf
od_SI01leaf
od_SI02leaf
od_SI03leaf
```
