---
title: "Lodes Plots"
author: "Owen Hart"
date: "7/15/2020"
output: html_document
---

```{r}
library(R.utils)
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(sf)
```

```{r}
lodes17 <- read_csv("../data/lodes/agg/jobs_all_2017.csv")

lodes16 <- read_csv("../data/lodes/agg/jobs_all_2016.csv")

lodes15 <- read_csv("../data/lodes/agg/jobs_all_2015.csv")




lodes17$h_state <- substr(lodes17$h_geocode_tract, 1, 2)
lodes17$h_county <- substr(lodes17$h_geocode_tract, 3, 5)
lodes17$h_tract <- substr(lodes17$h_geocode_tract, 6, 11)



lodes16$h_state <- substr(lodes16$h_geocode_tract, 1, 2)
lodes16$h_county <- substr(lodes16$h_geocode_tract, 3, 5)
lodes16$h_tract <- substr(lodes16$h_geocode_tract, 6, 11)


lodes15$h_state <- substr(lodes15$h_geocode_tract, 1, 2)
lodes15$h_county <- substr(lodes15$h_geocode_tract, 3, 5)
lodes15$h_tract <- substr(lodes15$h_geocode_tract, 6, 11)

#now work geoid


lodes17$w_state <- substr(lodes17$w_geocode_tract, 1, 2)
lodes17$w_county <- substr(lodes17$w_geocode_tract, 3, 5)
lodes17$w_tract <- substr(lodes17$w_geocode_tract, 6, 11)



lodes16$w_state <- substr(lodes16$w_geocode_tract, 1, 2)
lodes16$w_county <- substr(lodes16$w_geocode_tract, 3, 5)
lodes16$w_tract <- substr(lodes16$w_geocode_tract, 6, 11)


lodes15$w_state <- substr(lodes15$w_geocode_tract, 1, 2)
lodes15$w_county <- substr(lodes15$w_geocode_tract, 3, 5)
lodes15$w_tract <- substr(lodes15$w_geocode_tract, 6, 11)
```

Must load in Oregon XWalk csv
```{r}

#download.file(
#  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/or_xwalk.csv.gz", 
#  destfile = "../data/lodes/or_xwalk.csv.gz")
#gunzip("../data/lodes/or_xwalk.csv.gz")

or_xw <- read_csv("../data/lodes/or_xwalk.csv", col_types = cols(cty = col_character()))
```
What % of jobs flow from outside of Fairfax County?
```{r}
bool_cond17 <-lodes17$h_county != "065" & (lodes17$w_county == "065" & lodes17$w_state == "41")

bool_cond16 <-lodes16$h_county != "065" & (lodes16$w_county == "065" & lodes16$w_state == "41")

bool_cond15 <-lodes15$h_county != "065" & (lodes15$w_county == "065" & lodes15$w_state == "41")


flows_in_17 <-lodes17[bool_cond17, ]
flows_in_16 <-lodes16[bool_cond16, ]
flows_in_15 <-lodes15[bool_cond15, ]


pie(c(mean(bool_cond17) * 100, (1 - mean(bool_cond17)) * 100), labels = c("% Jobs in Wasco County\n flowing from outside (2017)", "% Jobs in Wasco County\n flowing from within (2017)"))

pie(c(mean(bool_cond16) * 100, (1 - mean(bool_cond16)) * 100), labels = c("% Jobs in Wasco County\n flowing from outside (2016)", "% Jobs in Wasco County\n flowing from within (2016)"))

pie(c(mean(bool_cond15) * 100, (1 - mean(bool_cond15)) * 100), labels = c("% Jobs in Wasco County\n flowing from outside (2015)", "% Jobs in Wasco County\n flowing from within (2015)"))
```


Of the flows into Wasco County from outside, what percent are from other counties in Oregon?
```{r}
#2017
perc_flows_othercnt_2017 <- mean(flows_in_17$h_state == "41") *100
agg_cnty_OR_2017 <- flows_in_17[flows_in_17$h_state == "41", ] %>%
  group_by(h_county) %>%
  summarise(S000 = sum(S000))
#Use OR crosswalk to get county names
or_xw$h_county <- substring(or_xw$cty, 3, 6)
essential <- or_xw %>%
  select(h_county, ctyname)
agg_cnty_OR_with_names_2017 <- inner_join(x = agg_cnty_OR_2017, y = unique(essential), by = "h_county")
trimmed_2017 <- agg_cnty_OR_with_names_2017[order(-agg_cnty_OR_with_names_2017$S000), ][1:10, ]
trimmed_2017 %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "dark blue", alpha=.7) +
  ggtitle("Number of Jobs flowing into Wasco County\n from other counties in Oregon (2017)") +
  xlab("Number of Jobs") + 
  ylab("County Name")
ggsave(path = "../output/lodes_maps/flows/in_flows", device = "png", filename = "in_county_flows_2017.png", plot = last_plot())

#2016
perc_flows_othercnt_2016 <- mean(flows_in_16$h_state == "41") *100
agg_cnty_OR_2016 <- flows_in_16[flows_in_16$h_state == "41", ] %>%
  group_by(h_county) %>%
  summarise(S000 = sum(S000))
agg_cnty_OR_with_names_2016 <- inner_join(x = agg_cnty_OR_2016, y = unique(essential), by = "h_county")
trimmed_2016 <- agg_cnty_OR_with_names_2016[order(-agg_cnty_OR_with_names_2016$S000), ][1:10, ]
trimmed_2016 %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "dark blue", alpha=.7) +
  ggtitle("Number of Jobs flowing into Wasco County\n from other counties in Oregon (2016)") +
  xlab("Number of Jobs") + 
  ylab("County Name")
ggsave(path = "../output/lodes_maps/flows/in_flows", device = "png", filename = "in_county_flows_2016.png", plot = last_plot())


#2015
perc_flows_othercnt_2015 <- mean(flows_in_15$h_state == "41") *100
agg_cnty_OR_2015 <- flows_in_15[flows_in_15$h_state == "41", ] %>%
  group_by(h_county) %>%
  summarise(S000 = sum(S000))
agg_cnty_OR_with_names_2015 <- inner_join(x = agg_cnty_OR_2015, y = unique(essential), by = "h_county")
trimmed_2015 <- agg_cnty_OR_with_names_2015[order(-agg_cnty_OR_with_names_2015$S000), ][1:10, ]
trimmed_2015 %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "dark blue", alpha=.7) +
  ggtitle("Number of Jobs flowing into Wasco County\n from other counties in Oregon (2015)") +
  xlab("Number of Jobs") + 
  ylab("County Name")
ggsave(path = "../output/lodes_maps/flows/in_flows", device = "png", filename = "in_county_flows_2015.png", plot = last_plot())

```


Outflows (those who live in wasco county but do not work there, open to other states)
```{r}
bool_cond17 <- lodes17$h_county == "065" & lodes17$w_county != "065"
bool_cond16 <- lodes16$h_county == "065" & lodes16$w_county != "065"
bool_cond15 <- lodes15$h_county == "065" & lodes15$w_county != "065"

flows_out_17 <-lodes17[bool_cond17, ]
flows_out_16 <-lodes16[bool_cond16, ]
flows_out_15 <-lodes15[bool_cond15, ]

pie(c(mean(bool_cond17) * 100, (1 - mean(bool_cond17)) * 100), labels = c("% Residents in Wasco County\n flowing to outside the county for work (2017)", "% Residents in Wasco County\n that work in Wasco County (2017)"))

pie(c(mean(bool_cond16) * 100, (1 - mean(bool_cond16)) * 100), labels = c("% Residents in Wasco County\n flowing to outside the county for work (2016)", "% Residents in Wasco County\n that work in Wasco County (2016)"))

pie(c(mean(bool_cond15) * 100, (1 - mean(bool_cond15)) * 100), labels = c("% Residents in Wasco County\n flowing to outside the county for work (2015)", "% Residents in Wasco County\n that work in Wasco County (2015)"))
```

Most popular counties that those who live in Wasco County but not work in, go to work
```{r}
out_agg_2017 <- flows_out_17 %>%
  group_by(w_county) %>%
  summarise(S000 = sum(S000))
or_xw$w_county <- substring(or_xw$cty, 3, 6)
essential <- or_xw %>%
  select(w_county, ctyname)
agg_out_names_2017 <- inner_join(x = out_agg_2017, y = unique(essential), by = "w_county")
trimmed_out_2017 <- agg_out_names_2017[order(-agg_out_names_2017$S000), ][1:10, ]
trimmed_out_2017 %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "red", alpha=.3) +
  ggtitle("Number of Jobs flowing out of Wasco County\n to other counties in Oregon (2017)") +
  xlab("Number of Jobs") + 
  ylab("County Name")
ggsave(path = "../output/lodes_maps/flows/out_flows", device = "png", filename = "out_county_flows_2017.png", plot = last_plot())

out_agg_2016 <- flows_out_16 %>%
  group_by(w_county) %>%
  summarise(S000 = sum(S000))
agg_out_names_2016 <- inner_join(x = out_agg_2016, y = unique(essential), by = "w_county")
trimmed_out_2016 <- agg_out_names_2016[order(-agg_out_names_2016$S000), ][1:10, ]
trimmed_out_2016 %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "red", alpha=.3) +
  ggtitle("Number of Jobs flowing out of Wasco County\n to other counties in Oregon (2016)") +
  xlab("Number of Jobs") + 
  ylab("County Name")
ggsave(path = "../output/lodes_maps/flows/out_flows", device = "png", filename = "out_county_flows_2016.png", plot = last_plot())

out_agg_2015 <- flows_out_15 %>%
  group_by(w_county) %>%
  summarise(S000 = sum(S000))
agg_out_names_2015 <- inner_join(x = out_agg_2015, y = unique(essential), by = "w_county")
trimmed_out_2015 <- agg_out_names_2015[order(-agg_out_names_2015$S000), ][1:10, ]
trimmed_out_2015 %>%
  mutate(ctyname = fct_reorder(ctyname, S000)) %>%
  ggplot(aes(S000, ctyname)) +
  geom_bar(stat="identity", fill= "red", alpha=.3) +
  ggtitle("Number of Jobs flowing out of Wasco County\n to other counties in Oregon (2015)") +
  xlab("Number of Jobs") + 
  ylab("County Name")
ggsave(path = "../output/lodes_maps/flows/out_flows", device = "png", filename = "out_county_flows_2015.png", plot = last_plot())
```







Mapping S. Wasco
```{r}
library(tigris)

wasco_points <- (blocks("OR", county = "Wasco"))
wasco_lines <- data.frame(wasco_points)

wasco_county <- st_read("../data/shps/county")
south_wasco_points <- st_read("../data/shps/swsd")

```

Reading in files
```{r}
main17 <- fread("../data/lodes/or_od_main_JT00_2017.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))

aux17 <- fread("../data/lodes/or_od_aux_JT00_2017.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))

full_data17 <- rbind(main17, aux17)


main16 <- fread("../data/lodes/or_od_main_JT00_2016.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))

aux16 <- fread("../data/lodes/or_od_aux_JT00_2016.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))

full_data16 <- rbind(main16, aux16)


main15 <- fread("../data/lodes/or_od_main_JT00_2015.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))

aux15 <- fread("../data/lodes/or_od_aux_JT00_2015.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))

full_data15 <- rbind(main15, aux15)
```

Filtering so that it only includes Wasco County (need to address S. Wasco part)
```{r}
wasco17 <- full_data17 %>% 
  filter(w_geocode %in% full_data17$w_geocode[substr(full_data17$w_geocode, 1, 5) == "41065"]|h_geocode %in% full_data17$h_geocode[substr(full_data17$h_geocode, 1, 5) == "41065"])


wasco16 <- full_data16 %>% 
  filter(w_geocode %in% full_data16$w_geocode[substr(full_data16$w_geocode, 1, 5) == "41065"]|h_geocode %in% full_data16$h_geocode[substr(full_data16$h_geocode, 1, 5) == "41065"])

wasco15 <- full_data15 %>% 
  filter(w_geocode %in% full_data15$w_geocode[substr(full_data15$w_geocode, 1, 5) == "41065"]|h_geocode %in% full_data15$h_geocode[substr(full_data15$h_geocode, 1, 5) == "41065"])
```

Joining with Wasco Lines
```{r}
wascolines_17 <- inner_join(wasco17, wasco_lines, by = c("w_geocode" = "GEOID10"))
wascolines_16 <- inner_join(wasco16, wasco_lines, by = c("w_geocode" = "GEOID10"))
wascolines_15 <- inner_join(wasco15, wasco_lines, by = c("w_geocode" = "GEOID10"))
wascolines_17
```


Grouping by work geocode (because that is going to be Wasco County)
```{r}
agg_17 <- wascolines_17 %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10, geometry) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))

agg_16 <- wascolines_16 %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10, geometry) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))

agg_15 <- wascolines_15 %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10, geometry) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
```


```{r}
agg_17
data.frame(south_wasco_points)

ggplot(south_wasco_points) +
  geom_sf() +
  geom_sf(data=agg_17[agg_17$S000 < 600, ], aes(fill= S000, geometry = geometry)) + #got rid of 1 outliar of 774
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in South Wasco County\n by Census Block in 2017") +
  labs(fill = "Number of Jobs")
```



2017 maps
```{r}
alljobs_2017 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_17[agg_17$S000 < 600, ], aes(fill= S000, geometry = geometry)) + #got rid of 1 outliar of 774
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in South Wasco County\n by Census Block in 2017") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2017", device = "png", filename = "alljobs_2017.png", plot = alljobs_2017)

goodsjobs_2017 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_17, aes(fill=SI01, geometry = geometry)) + 
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in Goods Producing industry sectors\n in South Wasco County by Census Block in 2017") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2017", device = "png", filename = "goodsjobs_2017.png", plot = goodsjobs_2017)



utilsjobs_2017 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_17, aes(fill=SI02, geometry = geometry)) + 
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in Trade, Transportation,\n and Utilities industry sectors\n in South Wasco County by Census Block in 2017") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2017", device = "png", filename = "utilsjobs_2017.png", plot = utilsjobs_2017)
  
servicejobs_2017 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_17[agg_17$SI03 < 600, ], aes(fill=SI03, geometry = geometry)) +  # 1 outliar
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in All Other Services industry sectors\n in South Wasco County by Census Block in 2017") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2017", device = "png", filename = "servicejobs_2017.png", plot = servicejobs_2017)

```

2016 maps
```{r}
alljobs_2016 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_16[agg_16$S000 < 600, ], aes(fill=S000, geometry = geometry)) + # 1 outliar
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in South Wasco County\n by Census Block in 2016") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2016", device = "png", filename = "alljobs_2016.png", plot = alljobs_2016)

goodsjobs_2016 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_16, aes(fill=SI01, geometry = geometry)) + 
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in Goods Producing industry sectors\n in South Wasco County by Census Block in 2016") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2016", device = "png", filename = "goodsjobs_2016.png", plot = goodsjobs_2016)


utilsjobs_2016 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_16, aes(fill=SI02, geometry = geometry)) + 
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in Trade, Transportation,\n and Utilities industry sectors\n in South Wasco County by Census Block in 2016") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2016", device = "png", filename = "utilsjobs_2016.png", plot = utilsjobs_2016)
  
servicejobs_2016 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_16[agg_16$SI03 < 600, ], aes(fill=SI03, geometry = geometry)) +  # 1 outliar
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in All Other Services industry sectors\n in South Wasco County by Census Block in 2016") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2016", device = "png", filename = "servicejobs_2016.png", plot = servicejobs_2016)
```

2015 maps
```{r}
alljobs_2015 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_15[agg_15$S000 < 600, ], aes(fill=S000, geometry = geometry)) + #1 outliar
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in South Wasco County\n by Census Block in 2015") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2015", device = "png", filename = "alljobs_2015.png", plot = alljobs_2015)

goodsjobs_2015 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_15, aes(fill=SI01, geometry = geometry)) + 
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in Goods Producing industry sectors\n in South Wasco County by Census Block in 2015") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2015", device = "png", filename = "goodsjobs_2015.png", plot = goodsjobs_2015)


utilsjobs_2015 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_15, aes(fill=SI02, geometry = geometry)) + 
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in Trade, Transportation,\n and Utilities industry sectors\n in South Wasco County by Census Block in 2015") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2015", device = "png", filename = "utilsjobs_2015.png", plot = utilsjobs_2015)
  
servicejobs_2015 <- ggplot(wasco_county) +
  geom_sf() +
  geom_sf(data = south_wasco_points, color = "red") +
  geom_sf(data=agg_15[agg_15$SI03 < 600, ], aes(fill=SI03, geometry = geometry)) +  # 1 outliar
  scale_fill_gradient(low="light gray", high="green") +
  ggtitle("Total number of jobs in All Other Services industry sectors\n in South Wasco County by Census Block in 2015") +
  labs(fill = "Number of Jobs")
ggsave(path = "../output/lodes_maps/2015", device = "png", filename = "servicejobs_2015.png", plot = servicejobs_2015)
```
```{r}
mapview(st_as_sf(agg_17), map.types = 'CartoDB.Positron',
          values = agg_17$SI01,
          at = quantile(agg_17$S000),
          label = agg_17$SI01, legend = TRUE,
          layer.name = 'South Wasco Goods Producing Sector\n Job Density by Census Block (2017)',
          zcol = 'SI01',
          col.regions = uvatheme(agg_17$S000))
```

interactive maps
```{r}
library(leaflet)
library(mapview)
uvatheme <- colorRampPalette(c('#D1E0BF', '#E57200'))


map_2017_data <- mapview(south_wasco_points, label = NULL, 
        alpha.region = 0, legend = FALSE,
        color = "dark green", popup = NULL,
        layer.name = "South Wasco School District") + 
  mapview(st_as_sf(agg_17), map.types = 'CartoDB.Positron',
          values = agg_17$S000,
          at = quantile(agg_17$S000),
          label = agg_17$S000, legend = TRUE,
          layer.name = 'South Wasco Job Density by Census Block (2017)',
          zcol = 'S000',
          col.regions = uvatheme(agg_17$S000)) +
  
  mapview(st_as_sf(agg_17), map.types = 'CartoDB.Positron',
          values = agg_17$SI01,
          at = quantile(agg_17$S000),
          label = agg_17$SI01, legend = TRUE,
          layer.name = 'South Wasco Goods Producing Sector\n Job Density by Census Block (2017)',
          zcol = 'SI01',
          col.regions = uvatheme(agg_17$S000)) +
  
  mapview(st_as_sf(agg_17), map.types = 'CartoDB.Positron',
          values = agg_17$SI02,
          at = quantile(agg_17$S000),
          label = agg_17$SI02, legend = TRUE,
          layer.name = 'South Wasco Trade, Transportation,and Utilities\n Job Density by Census Block
          (2017)',
          zcol = 'SI02',
          col.regions = uvatheme(agg_17$S000)) +
  
  mapview(st_as_sf(agg_17), map.types = 'CartoDB.Positron',
          values = agg_17$SI03,
          at = quantile(agg_17$S000),
          label = agg_17$SI03, legend = TRUE,
          layer.name = 'South Wasco All Other Services industry\n Job Density by Census Block
          (2017)',
          zcol = 'SI02',
          col.regions = uvatheme(agg_17$S000))





map_2017_data
```

make it leaflet
```{r}
map2017leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 1,
    opacity = 1,
    group = "Basemap")


qtileS000 <- colorQuantile(c('#D1E0BF', '#E57200'), agg_17$S000, 5)

map2017leaf <- map2017leaf %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "All Jobs",
    fillColor = ~qtileS000(agg_17$S000),
    label = agg_17$S000)
map2017leaf <- map2017leaf %>%
      addLegend(
        data = agg_17,
        "bottomright",
        pal = qtileS000,
        values = ~ S000,
        title = "Wasco County Job Density",
        opacity = 1,
        group = "All Jobs",
        na.label = "NA")

#SI01
map2017leaf <- map2017leaf %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "Goods Producing Sector",
    fillColor = ~qtileS000(agg_17$SI01),
    label = agg_17$SI01) %>%
  addLegend(
    data = agg_17,
    "bottomright",
    pal = qtileS000,
    values = ~ SI01,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "Goods Producing Sector",
    na.label = "NA")

#SI02
map2017leaf <- map2017leaf %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "Trade, Transportation,and Utilities Sector",
    fillColor = ~qtileS000(agg_17$SI02),
    label = agg_17$SI02) %>%
  addLegend(
    data = agg_17,
    "bottomright",
    pal = qtileS000,
    values = ~ SI02,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "Trade, Transportation,and Utilities Sector",
    na.label = "NA")

#SI03
map2017leaf <- map2017leaf %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "All Other Services Sector",
    fillColor = ~qtileS000(agg_17$SI03),
    label = agg_17$SI03) %>%
  addLegend(
    data = agg_17,
    "bottomright",
    pal = qtileS000,
    values = ~ SI03,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "All Other Services Sector",
    na.label = "NA")

map2017leaf <- map2017leaf %>%
      addLayersControl(
        baseGroups = c("South Wasco School District"),
        overlayGroups = c("All Jobs", "Goods Producing Sector", "Trade, Transportation,and Utilities Sector", "All Other Services Sector"),
        options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("Goods Producing Sector", "Trade, Transportation,and Utilities Sector", "All Other Services Sector"))
```


2016
```{r}
map2016leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 1,
    opacity = 1,
    group = "Basemap")


qtileS000 <- colorQuantile(c('#D1E0BF', '#E57200'), agg_16$S000, 5)

map2016leaf <- map2016leaf %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "All Jobs",
    fillColor = ~qtileS000(agg_16$S000),
    label = agg_16$S000)
map2016leaf <- map2016leaf %>%
      addLegend(
        data = agg_16,
        "bottomright",
        pal = qtileS000,
        values = ~ S000,
        title = "Wasco County Job Density",
        opacity = 1,
        group = "All Jobs",
        na.label = "NA")

#SI01
map2016leaf <- map2016leaf %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "Goods Producing Sector",
    fillColor = ~qtileS000(agg_16$SI01),
    label = agg_16$SI01) %>%
  addLegend(
    data = agg_16,
    "bottomright",
    pal = qtileS000,
    values = ~ SI01,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "Goods Producing Sector",
    na.label = "NA")

#SI02
map2016leaf <- map2016leaf %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "Trade, Transportation,and Utilities Sector",
    fillColor = ~qtileS000(agg_16$SI02),
    label = agg_16$SI02) %>%
  addLegend(
    data = agg_16,
    "bottomright",
    pal = qtileS000,
    values = ~ SI02,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "Trade, Transportation,and Utilities Sector",
    na.label = "NA")

#SI03
map2016leaf <- map2016leaf %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "All Other Services Sector",
    fillColor = ~qtileS000(agg_16$SI03),
    label = agg_16$SI03) %>%
  addLegend(
    data = agg_16,
    "bottomright",
    pal = qtileS000,
    values = ~ SI03,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "All Other Services Sector",
    na.label = "NA")

map2016leaf <- map2016leaf %>%
      addLayersControl(
        baseGroups = c("South Wasco School District"),
        overlayGroups = c("All Jobs", "Goods Producing Sector", "Trade, Transportation,and Utilities Sector", "All Other Services Sector"),
        options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("Goods Producing Sector", "Trade, Transportation,and Utilities Sector", "All Other Services Sector"))
```

2015
```{r}
map2015leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 1,
    opacity = 1,
    group = "Basemap")


qtileS000 <- colorQuantile(c('#D1E0BF', '#E57200'), agg_15$S000, 5)

map2015leaf <- map2015leaf %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "All Jobs",
    fillColor = ~qtileS000(agg_15$S000),
    label = agg_15$S000)
map2015leaf <- map2015leaf %>%
      addLegend(
        data = agg_15,
        "bottomright",
        pal = qtileS000,
        values = ~ S000,
        title = "Wasco County Job Density",
        opacity = 1,
        group = "All Jobs",
        na.label = "NA")

#SI01
map2015leaf <- map2015leaf %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "Goods Producing Sector",
    fillColor = ~qtileS000(agg_15$SI01),
    label = agg_15$SI01) %>%
  addLegend(
    data = agg_15,
    "bottomright",
    pal = qtileS000,
    values = ~ SI01,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "Goods Producing Sector",
    na.label = "NA")

#SI02
map2015leaf <- map2015leaf %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "Trade, Transportation,and Utilities Sector",
    fillColor = ~qtileS000(agg_15$SI02),
    label = agg_15$SI02) %>%
  addLegend(
    data = agg_15,
    "bottomright",
    pal = qtileS000,
    values = ~ SI02,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "Trade, Transportation,and Utilities Sector",
    na.label = "NA")

#SI03
map2015leaf <- map2015leaf %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = 1,
    group = "All Other Services Sector",
    fillColor = ~qtileS000(agg_15$SI03),
    label = agg_15$SI03) %>%
  addLegend(
    data = agg_15,
    "bottomright",
    pal = qtileS000,
    values = ~ SI03,
    title = "Wasco County Job Density",
    opacity = 1,
    group = "All Other Services Sector",
    na.label = "NA")

map2015leaf <- map2015leaf %>%
      addLayersControl(
        baseGroups = c("South Wasco School District"),
        overlayGroups = c("All Jobs", "Goods Producing Sector", "Trade, Transportation,and Utilities Sector", "All Other Services Sector"),
        options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("Goods Producing Sector", "Trade, Transportation,and Utilities Sector", "All Other Services Sector"))
```

Showing all maps
```{r}
map2017leaf
map2016leaf
map2015leaf
```


















raster
```{r}
library(raster)
library(stars)
#install.packages('raster', repos = 'http://r-forge.r-project.org/', type = 'source')
raster::raster("../data/lodes/raster/CDL_2017_41065.tif")

```



```{r}
#raster: install.packages("raster")
rgdal: install.packages("rgdal")
library(raster)

r <- raster("../data/lodes/raster/CDL_2017_41065.tif")
#map2017leaf %>%
#  addRasterImage()
```

